Geometric variational problems

### 2020-11-09 NSF2020/Description/preliminarywork.tex
yny

\begin{figure}[!]
\begin{center}
\includegraphics[width=1\textwidth]{figures/PW_fig1A-D.pdf}
\includegraphics[width=1\textwidth]{figures/PW_fig1E-H.pdf}
\end{center}\vspace{-0.3in}
\caption{(A) A vesicle formed by amphiphilic particles in shear flow,
  and the tank-treading motion (B)--(D). The separation of particle
  pairs in (B) and (C) illustrate inter-leaflet slip.  (E)--(G)
  Tank-treading reaches a steady state in elliptical aspect ratio,
  major-axis angle, and circumference.}
\label{fig:tanktreading}
\end{figure}
\section{Preliminary Work on Two-dimensional Vesicle Hydrodynamics in Shear Flow} 
%The motion of vesicles in shear flow is an important
%problem in the applied mathematics because simulations can reveal mechanical
%properties of membranes and lead to an enhanced understanding of 
%deformable particle laden flows \cite{Sinha15}. 
%
Over the past decades, researchers have used a number of mathematical tools 
to model vesicles in shear flow, including  phase field \cite{DuLiuWang2004_JCP,BibenKassnerMisbah2005_PRE}, 
level set \cite{DoyeuxGuyotChabannesEtAl2013_JCAM}, boundary integral \cite{Shravan09,Rahimian15} and 
immersed boundary approaches \cite{KimLai2010_JCP,KimLai2012_PRE,HuLaiSeolEtAl2016_JCP}. All these approaches
assume a mathematical surface, whether implicitly or explicitly, 
and define an elastic bending energy of the surface. The vesicle obeys fluid
transport and in turn the fluid balances shear stress with the vesicle's bending force.

@@

To implement a vesicle in shear flow in the context of hydrophobic potentials and mobility problem, 
we consider a shear flow $\mathbf{u}_{\infty} = Uy\mathbf{i}_x$ in the direction of the $x$-axis
(Figure \ref{fig:tanktreading}A).
This field satisfies the linear Stokes system but does not give rise to a rigid motion at the particle interfaces. 
To have a rigid motion, we change variables $\mathbf{u} = \tilde{\mathbf{u}}+ \mathbf{u}_{\infty}$ and 
for the new field $\tilde{\mathbf{u}}$ vanishing at infinity we let 
$\tilde{\mathbf{u}}|_{\partial P_i} = \mathbf{v}_i + \boldsymbol{\omega}_i \times (\mathbf{x} - \mathbf{a}_i)$ 
where $(\mathbf{v}_i,\boldsymbol{\omega}_i)$ are the unknown translation and angular velocities of the 
$i$th particle $P_i.$

\begin{wrapfigure}[17]{l}{0.4\textwidth}
\centerline{\includegraphics[width=0.4\textwidth]{figures/PW_fig2.pdf}}
\caption{\label{fig:demixing} An initial assembly of small and 
large particles spontaneous segregates into two smaller bodies. }
\end{wrapfigure}
The HAP simulations show vesicle tank-treading. Under the external shear flow, the initially circular 
vesicle rotates in the clockwise direction. As the rate of rotation increases, the vesicle approaches
a steadily tank-treading ellipse. In Figure \ref{fig:tanktreading}B-D, the solid curves are ellipses fit to the particle centers
and midplane respectively. In the non-dimensionalized system, the particles have diameter 2, on the order of $\rho,$ 
and the vesicle diameter is about 14. \todo[inline]{missing units. nm?}
Figure~\ref{fig:tanktreading}E shows the aspect ratio of the major to minor axes reaching an equilibrium value in the 
red and blue curves, yet oscillating in the high-shear rate (yellow) curve.
The tank-treading vesicle elongates and becomes more horizontal 
with an increase in flow rate or 
with a decrease in stiffness (effected by decreasing $\rho = 4$ to $\rho = 2$).

@@

\begin{wrapfigure}[12]{r}{0.2\textwidth}
\centerline{\includegraphics[width=0.2\textwidth]{figures/PW_fig5.pdf}}
\caption{\label{fig:rupture} Rupture of tank-treading vesicle under strong shear flow.}
\end{wrapfigure}
There are a number of advantages to using our particle based approach.
For one, the HAP theory automatically accounts for the existence of multiple phases. 
For example, we can vary lipid length, spontaneous curvature and bending rigidity 
by introducing differentiating particle shapes and hydrophobic boundary conditions (Figure \ref{fig:demixing}).
Continuum theory deals with multiple phases through additional surface densities that must
then satisfy specialized transport equations \cite{Lowengrub07,MikuckiZhou17}. 
As illustrated in Figure \ref{fig:tanktreading}, the particle
based approach supports inter-leaflet slip, and this can be used to determine inter-leaflet
and in-plane shear viscosities.

The ability to form discontinuities is perhaps the HAP method's greatest 
strength. Indeed, a strong, extensional shear flow must cause membrane rupture since the 
energy of stretching must exceed the pore nucleation and widening barriers 
for sufficiently large flow strengths. 
Some disadvantages are that the hydrophobic interaction does not constrain vesicle volume. 
Instead, changes in volume are rate limited by inter-particle spacing,
which may be undesirable in case of a mathematically strict volume constraint.
% set vesicle volume, as is done in the study of liposomes
%using auxiliary constraint equations for instance. 
%Also, the boundary integral method formulation
%relies on linearity of the Stokes equations. There has been some progress in
%boundary integral methods for the non-linear Navier-Stokes equations [ref].
%We point out that some non-Newtonian effects in polymers are a consequence of hydrophobic
%and steric molecular interactions like the ones presented in this proposal.

In recent years, researchers have developed a number of numerical methods for calculating
energy minimizing steady equilibrium shapes of lipid bilayer membranes, vesicles and red blood cells.
These approaches range from the finite element method~\cite{Bartels,Peng13,RyKlYaCo16,Sinha15}, 
phase field method \cite{Du05,QiangDu08,Lowengrub13} to immersed
boundary method  \cite{Hu,Hu13, KimLai2010_JCP}.
PI Ryham and collaborators led in part the development of phase field
functionals of membrane elastic energy and approaches to coupling membrane elasticity to fluids  \cite{0951-7715-18-3-016,Du05,DuEuler,QiangDu09}.

>>>>

\begin{figure}[!]
\begin{center}
\includegraphics[width=1\textwidth]{figures/PW_fig1A-D.pdf}
\includegraphics[width=1\textwidth]{figures/PW_fig1E-H.pdf}
\end{center}\vspace{-0.3in}
\caption{(A) A vesicle formed by amphiphilic particles in shear flow,
  and the tank-treading motion (B)--(D). The separation of particle
  pairs in (B) and (C) illustrate inter-leaflet slip.  (E)--(G)
  Tank-treading reaches a steady state in elliptical aspect ratio,
  major-axis angle, and circumference.}
\label{fig:tanktreading}
\end{figure}
\section{Preliminary Work on Two-dimensional Vesicle Hydrodynamics in Shear Flow} 
%The motion of vesicles in shear flow is an important
%problem in the applied mathematics because simulations can reveal mechanical
%properties of membranes and lead to an enhanced understanding of 
%deformable particle laden flows \cite{Sinha15}. 
%
Over the past decades, researchers have used a number of mathematical tools 
to simulate vesicles in a shear flow, including  lattice Boltzmann \cite{KaouiHartingMisbah2011_PRE}, coarse-grained Brownian dynamics \cite{NoguchiTakasu2002_BJ},
phase field \cite{DuLiuWang2004_JCP,BibenKassnerMisbah2005_PRE}, 
level set \cite{DoyeuxGuyotChabannesEtAl2013_JCAM}, boundary integral \cite{Shravan09,Rahimian15} and 
immersed boundary approaches \cite{KimLai2010_JCP,KimLai2012_PRE,HuLaiSeolEtAl2016_JCP}. Most of these approaches
assume a mathematical surface, whether implicitly or explicitly, 
and define an elastic bending energy of the surface. 
%The vesicle obeys fluid
%transport and in turn the fluid balances shear stress with the vesicle's bending force.

@@

To implement a vesicle in shear flow in the context of hydrophobic potentials and mobility problem, 
we consider a shear flow in the far-field $\mathbf{u}_{\infty} = Uy\mathbf{i}_x$ in the direction of the $x$-axis
(Figure \ref{fig:tanktreading}A).
%
%This field satisfies the linear Stokes system but does not give rise to a rigid motion at the particle interfaces. 
%To have a rigid motion, we change variables $\mathbf{u} = \tilde{\mathbf{u}}+ \mathbf{u}_{\infty}$ and 
%for the new field $\tilde{\mathbf{u}}$ vanishing at infinity we let 
%$\tilde{\mathbf{u}}|_{\partial P_i} = \mathbf{v}_i + \boldsymbol{\omega}_i \times (\mathbf{x} - \mathbf{a}_i)$ 
%where $(\mathbf{v}_i,\boldsymbol{\omega}_i)$ are the unknown translation and angular velocities of the 
%$i$th particle $P_i.$  
%
%\begin{wrapfigure}[17]{l}{0.4\textwidth}
%\centerline{\includegraphics[width=0.4\textwidth]{figures/PW_fig2.pdf}}
%\caption{\label{fig:demixing} An initial assembly of small and 
%large particles spontaneous segregates into two smaller bodies. }
%\end{wrapfigure}
The HAP simulations show vesicle tank-treading. Under the external shear flow, the initially circular 
vesicle rotates in the clockwise direction. As the rate of rotation increases, the vesicle approaches
a steadily tank-treading ellipse. In Figure \ref{fig:tanktreading}B-D, the solid curves are ellipses fit to the particle centers
and midplane respectively. In the non-dimensionalized system, the particles have diameter 2, on the order of $\rho,$ 
and the vesicle diameter is about 14. 
%\todo[inline]{missing units. nm?}
%
\begin{wrapfigure}[12]{r}{0.2\textwidth}
\centerline{\includegraphics[width=0.2\textwidth]{figures/PW_fig5.pdf}}
\caption{\label{fig:rupture} Rupture of tank-treading vesicle under strong shear flow.}
\end{wrapfigure}
%
Figure~\ref{fig:tanktreading}E shows the aspect ratio of the major to minor axes reaching an equilibrium value in the 
red and blue curves, yet oscillating in the high-shear rate (yellow) curve.
The tank-treading vesicle elongates and becomes more horizontal 
with an increase in flow rate or 
with a decrease in stiffness (effected by decreasing $\rho = 4$ to $\rho = 2$).

@@

%\begin{wrapfigure}[12]{r}{0.2\textwidth}
%\centerline{\includegraphics[width=0.2\textwidth]{figures/PW_fig5.pdf}}
%\caption{\label{fig:rupture} Rupture of tank-treading vesicle under strong shear flow.}
%\end{wrapfigure}
There are a number of advantages to using our particle based approach.
For one, the HAP theory automatically accounts for the existence of multiple phases. 
For example, we can vary lipid length, spontaneous curvature and bending rigidity 
by introducing differentiating particle shapes and hydrophobic boundary conditions (Figure \ref{fig:demixing}).
Continuum theory deals with multiple phases through additional surface densities that must
then satisfy specialized transport equations \cite{Lowengrub07,MikuckiZhou17}. 
As illustrated in Figure \ref{fig:tanktreading}, the particle
based approach supports inter-leaflet slip, and this can be used to determine inter-leaflet
and in-plane shear viscosities. 
\begin{wrapfigure}[17]{l}{0.4\textwidth}
\centerline{\includegraphics[width=0.4\textwidth]{figures/PW_fig2.pdf}}
\caption{\label{fig:demixing} An initial assembly of small and 
large particles spontaneous segregates into two smaller bodies. }
\end{wrapfigure}
%

The greatest strength of using HAP to model lipid bilayer membrane is the ability
to form discontinuities (interfacial singularity) from first physical principle without any artificial manipulation to rearrangement the interface.
%The ability to form discontinuities is perhaps the HAP method's greatest strength.
 Indeed, a strong hear flow must cause membrane rupture since the 
energy of stretching must exceed the pore nucleation and widening barriers 
for sufficiently large flow strengths. 
Some disadvantages are that the hydrophobic interaction does not constrain vesicle volume. 
Instead, changes in volume are rate limited by inter-particle spacing,
which may be undesirable in case of a mathematically strict volume constraint.
% set vesicle volume, as is done in the study of liposomes
%using auxiliary constraint equations for instance. 
%Also, the boundary integral method formulation
%relies on linearity of the Stokes equations. There has been some progress in
%boundary integral methods for the non-linear Navier-Stokes equations [ref].
%We point out that some non-Newtonian effects in polymers are a consequence of hydrophobic
%and steric molecular interactions like the ones presented in this proposal.

In recent years, researchers have developed a number of numerical methods for calculating
energy minimizing steady equilibrium shapes of lipid bilayer membranes, vesicles and red blood cells.
These approaches range from the finite element method~\cite{Bartels,Peng13,RyKlYaCo16,Sinha15}, 
phase field method \cite{Du05,QiangDu08,Lowengrub13} to immersed
boundary method  \cite{Hu,Hu13, KimLai2010_JCP}.
PI RR and collaborators led in part the development of phase field
functionals of membrane elastic energy and approaches to coupling membrane elasticity to fluids  \cite{0951-7715-18-3-016,Du05,DuEuler,QiangDu09}.

### 2020-11-09 NSF2020/Description/specificaim1_ver2.tex
yny

\section{Proposed Research}
\label{sec:proposed-work}
The proposed research develops fast,
high-order-accurate, parallel numerical algorithms for large-scale
simulations of the collective hydrodynamics of amphiphilic particles in a viscous solvent.
%
We have demonstrated that our hydrophobic attraction with repulsion potential (HARP) approach efficiently simulates
self-assembly of amphiphilic particles into two-dimensional micelles, bilayer membranes, and vesicles \cite{Fu2018_SIAM}, and
recreates the tank-treading phenomenon in external shear flows
\cite{Fu20} \todo[inline]{!!let's put a proceedings here so we have something to cite!!}.
%
While the results show great promise in the field of collective body hydrodynamics,
several outstanding issues need to be addressed. These include a thorough 
analysis of elastic properties of our coarse-grained bilayers, and 
efficiently simulating three-dimensional collective hydrodynamics of amphiphilic particles.
Finally, we must mathematically characterize the variational behavior of HARP under appropriate homogenized limits.

\subsection{Specific Aim 1: Collective elastic properties of amphiphilic particles}
\label{subsec:specific_aim_1}

%\begin{figure}
%\begin{center}
%\includegraphics[width=0.9\textwidth]{figures/SA1_fig1.pdf}
%\end{center}
%\caption{\footnotesize (A--C) The splay ($\Div \mathbf{d}$), 
%saddle splay ($\det \mathsf{D}$) and twist ($\Curl \mathbf{d}$) elastic distortions of 
%a monolayer. (D) The monolayer neutral surface $\Sigma$,  
%director $\mathbf{d}$ and unit normal $\mathbf{n}$ in local coordinates.
%(E--F) Lipids are able to tilt from  the surface normal, and stretch.
%(H--J) Intermediates of membrane fusion and (K) experimental image of a stalk \cite{Aeffner2012}. }
%\label{fig:distortions}
%\end{figure}
The goal of Specific Aim 1 is characterize the material properties of many-body, self-assembled amphiphiles.
For amphiphiles assembled into bilayers, these properties are described by membrane continuum mechanics.
The end-goal is to be able to map the parameters of the particle-based model onto the elastic moduli from continuum theory.
If we achieve this goal, then simulators can use the hydrophobic attraction force calculations
to model bilayers with specific composition. These calculations have provably less computational complexity than
those of molecular dynamics simulations and possess the molecular granularity lacking from continuum models.

>>>>

\section{Proposed Research}
\label{sec:proposed-work}
%The proposed research develops fast,
%high-order-accurate, parallel numerical algorithms for large-scale
%simulations of the collective hydrodynamics of amphiphilic particles in a viscous solvent.
%%
We have demonstrated that our hydrophobic attraction with repulsion potential (HARP) approach efficiently simulates
self-assembly of amphiphilic particles into two-dimensional micelles, bilayer membranes, and vesicles \cite{Fu2018_SIAM}, and
recreates the tank-treading phenomenon in external shear flows
\cite{Fu20}.
%\todo[inline]{!!let's put a proceedings here so we have something to cite!!}.
%
While the results show great promise in the field of collective body hydrodynamics,
several outstanding issues need to be addressed. These include a thorough 
analysis of elastic properties of our coarse-grained bilayers, and 
efficiently simulating three-dimensional collective hydrodynamics of amphiphilic particles.
Finally, we must mathematically characterize the variational behavior of HARP under appropriate homogenized limits.

\subsection{Specific Aim 1: Material properties of self-assembly of amphiphiles}
\label{subsec:specific_aim_1}

%\begin{figure}
%\begin{center}
%\includegraphics[width=0.9\textwidth]{figures/SA1_fig1.pdf}
%\end{center}
%\caption{\footnotesize (A--C) The splay ($\Div \mathbf{d}$), 
%saddle splay ($\det \mathsf{D}$) and twist ($\Curl \mathbf{d}$) elastic distortions of 
%a monolayer. (D) The monolayer neutral surface $\Sigma$,  
%director $\mathbf{d}$ and unit normal $\mathbf{n}$ in local coordinates.
%(E--F) Lipids are able to tilt from  the surface normal, and stretch.
%(H--J) Intermediates of membrane fusion and (K) experimental image of a stalk \cite{Aeffner2012}. }
%\label{fig:distortions}
%\end{figure}
The goal of Specific Aim 1 is characterize the material properties of many-body, self-assembled amphiphiles.
For amphiphiles assembled into bilayers, these properties are described by membrane continuum mechanics.
Our goal is to be able to map the parameters of the particle-based model onto the elastic moduli from continuum theory.
Results from this goal will facilitate simulators to use the hydrophobic attraction force calculations
to model bilayers with specific composition. These calculations have provably less computational complexity than
those of molecular dynamics simulations and possess the molecular granularity lacking from continuum models.
